{\Huge Supersymmetry} \\ \vspace{5mm} Lecture notes for FYS5190/FYS9190 Supersymmetry

### 2023-09-05 groups.tex
Finally completed all updates to chap. 1 for 2023. I hope...

\subsubsection{$SO(2)$}
Given that $SO(2)$ has only one parameter, we can write a general group member $R$ as parameterised by $\theta$
\[
R(\theta)=\left[\begin{matrix} \cos\theta &  -\sin\theta \\ \sin\theta&  \cos\theta \end{matrix}\right].
\]
As expected, we recognise this as the matrix of rotations of an angle $\theta$ around a point in the plane, and it represents the (orientation preserving) symmetries of a circle. Some tinkering with this representation will show that $SO(2)$ is in fact an abelian group, despite the matrix definition. From a physical viewpoint this should be expected: the order of rotations in the plane should not matter.

@@

A similar calculation for $SO(3)$, which also has three free parameters, gives the generators $X^i=\frac{1}{2}L^i$, where $L^i$ are the three matrices in Eq~(\ref{eq:SO3_generators}). These generators have the exact same commutator, $[X_i,X_j]=\epsilon_{ijk}X^k$, up to a constant factor, indicating a structural similarity between the groups $SU(2)$ and $SO(3)$. For $SO(3)$ we can further find a corresponding set of differential generators
\[
L_x=-i\left(y\frac{\partial}{\partial z}-z\frac{\partial}{\partial y} \right),\quad
L_y=-i\left(z\frac{\partial}{\partial x}-x\frac{\partial}{\partial z} \right),\quad
L_z=-i\left(x\frac{\partial}{\partial y}-y\frac{\partial}{\partial x} \right),
\]
see Ex.~\ref{ex:SO2_diff_generators}.
Up to a constant factor of $\hbar$ you may recognise these operators as the angular momentum operators in quantum mechanics. This is not a coincidence, the angular momentum operators in quantum mechanics are a representation of the generators of the group of rotations in three dimensions. In fact, the possibility of going back and forth between differential representations and matrix representations of the Lie group is the essence of the duality between the Schr\"odinger and Heisenberg pictures of quantum mechanics.

@@

@@

From what we learnt in Section \ref{sec:Lie_groups} the generators $X_i$ of an $d$-dimensional Lie group then span a $d$-dimensional Lie algebra with the commutator as the Lie bracket. Since this is also a vector space, any element $X$ of the algebra can be written in terms of the generators $X_i$, as $X=a_iX^i$. This is called the {\bf tangent space} of the group at the identity. The differences between choices in the definition of the generators and the normalisation of the structure constants disappear as they can be absorbed into a rescaling and linear combinations of the basis $\{X_i\}_{i=1}^d$, this makes the Lie algebra of the group {\it unique}.

@@

We saw in the previous subsection that -- just as for groups -- we have many different way of writing down a single Lie algebra. We then need a formal definition of representations of Lie algebras.
\df{A {\bf representation} $\rho$ {\bf of a Lie algebra} $\mathfrak l$ on a vector space $V$  is a map
\[\rho: \mathfrak{l} \to \mathfrak{gl}(V), \]
where $ \mathfrak{gl}$ is the Lie algebra of all linear maps $V\to V$ (endomorphisms) with its bracket defined as the composition $[r,s]\equiv r\circ s-s\circ r$, for all $r,s\in \mathfrak{gl}$, and where the map $\rho$ is homomorphic (structure preserving) for the Lie algebra bracket so that for all $x,y\in \mathfrak{l} $,
\[ \rho([x,y]) =\rho(x)\circ\rho(y)-\rho(y)\circ\rho(x). \]
}
Again this is an abstract definition that is actually simple than it looks. For finite dimensional vector spaces $\mathfrak{gl}$ is the Lie algebra of the corresponding general linear matrix group, just as the definition of representations of groups was a map to the general linear group $GL(n)$. This algebra of $GL(n)$, called $\mathfrak{gl}(n)$, can be written up in terms of $n\times n$ matrices and their commutators, and in this case the multiplication operation $\circ$ is just matrix multiplication in hiding. One word of warning here though, while the group $GL(n)$ contained invertible matrices, there is no guarantee that its generators are invertible matrices, {\it e.g.} we saw earlier that the generators of $SO(3)$ were the matrices $L_i$ in Eq~(\ref{eq:SO3_generators}) that are not invertible since $\det L_i=0$.

What makes this definition of representations even simpler for Lie algebras is that {\bf Ado's theorem}~\cite{Ado:1935} proves that every finite-dimensional Lie algebra has a faithful representation on a finite-dimensional vector space.\footnote{For a modern version of a proof of Ado's theorem, see Terence Tao's blog \url{https://terrytao.wordpress.com/2011/05/10/ados-theorem/}.} In other words, the Lie algebras for the groups we most care about, Lie groups with a finite number of parameters and thus finite number of generators, is guaranteed to have representations using matrices.

As for groups, we can talk about irreducible representations of the Lie algebras, which are representations such that $V$ has no proper invariant subspace under the representation, meaning that we cannot construct proper sub-representations.
% Maybe say something about similarity transforms between different Lie algebras

In Sec.~\ref{sec:generators} we saw the two sets of generators of the Lie algebras $\mathfrak{su(2)}$ and $\mathfrak{so(3)}$, derived from the defining matrix representations of the groups. Since this is the same algebra this shows two distinct representations of the same algebra.

%%%
\subsection{Exponential map}
\label{sec:expmap}
As we discussed in Section \ref{sec:Lie_groups}, the generators describe the local structure of the group near the identity element. We can now finally look at how the group (and matrix representation) is reconstructed from the algebra. For this we use what is called the {\bf exponential map}. 
\df{The {\bf exponential map} from a finite-dimensional Lie algebra $\mathfrak l$ of a group $G$ to $G$ is defined by $\exp:\mathfrak l\to G$, where for $X\in\mathfrak l$ we get the element $g\in G$ given by
\begin{equation}
g=\exp(iX)\equiv\sum_{n=0}^{\infty}\frac{(ia_iX^i)^n}{n!}.
\end{equation}
Here the $X_i$ are the generators of $G$ and thus members of the Lie algebra, and $X=a_iX^i$.
}
The infinite sum in this definition is nothing more than the formal series definition of the matrix exponential introduced in Sec.~\ref{sec:unitary_groups}. By Ado's theorem any (finite-dimensional) Lie algebra can be represented by matrices, so this construction is always valid, however, we can also use the exponential map for differential representations where the $X_i$ are differential operators, and for infinite-dimensional Lie groups, but the technical aspects why this extension works is a little beyond the scope of these notes.

@@

If we have two elements from the map $g=\exp{(iX)}$ and $g'=\exp{(iY)}$, then $gg'=\exp{(iX)}\exp{(iY)}$, and we would like to show that  $\exp{(iX)}\exp{(iY)}=\exp{(iZ)}$, where $Z\in\mathfrak l$. However, unless $X$ and $Y$ commute we cannot  easily join them in the same exponential. We here use the {\bf Baker--Campbell--Haussdorff  formula},
\begin{equation}
Z=X+Y+\frac{1}{2}[X,Y]+\frac{1}{12}[[X,Y],Y]-\frac{1}{12}[[X,Y],X]+\ldots,
\label{eq:BKH_matrices}
\end{equation}
where the omitted terms involve Lie brackets of four or more elements, and where every element in the series contains the commutator $[X,Y]$. If $X$ and $Y$ are close enough to the zero in $\mathfrak l$ (corresponding to the identity element in $G$) then this series can be shown to converge to an element in $\mathfrak l$.\footnote{This formula holds in general, even if the elements in the Lie algebra ${\mathfrak l}$ are not represented as finite-dimensional matrices.} Thus the exponential constructs a group at least locally around the identity element.

@@

For $SU(2)$ the generators taken from the fundamental representation $\mathbf 2$ were the Pauli matrices $\sigma_i$, and the exponential map is thus, as we alluded to earlier, $e^{i\alpha_i\sigma^i}$, where $\alpha_i$ are the parameters. Beyond the fundamental representations $SU(2)$ has irreducible representations in all dimensions $\mathbf 2$, $\mathbf 3$, $\mathbf 4$, $\mathbf 5$, \ldots.

@@

Note that the dimension $n$ of the matrices in the adjoint representation for $\mathfrak so(m)$ and $\mathfrak su(m)$, which is equal to the number of independent parameters, is $n=m(m-1)/2$ and $n=m^2-1$, respectively. For the matrices in the representation derived from the defining representation of the group the dimensions is the same as the dimension of the defining matrices, $m$. This means that in most cases the matrices in the adjoint representation is larger, with the exception of  $\mathfrak so(2)$ and  $\mathfrak so(3)$.

We can now briefly return to the $SU(2)$ and $SO(3)$ examples. These groups have the same structure constants because they have the same commutators between generators. Thus the adjoint representations of their algebras are the same. With structure constants $C^{~~k}_{ij}=\epsilon_{ij}^{~~k}$ we get the elements of the adjoint matrices $(M_i)_{jk} =-i\epsilon_{ijk}$, which gives $M_i=iL_i$, with the $L_i$ being the matrices in (\ref{eq:SO3_generators}).

From the adjoint representation of the Lie algebra we can construct  the adjoint representations of the Lie group. For example the adjoint representation of  $\mathfrak su(2)$ or $\mathfrak so(3)$ has the exponential map $\exp{(ia_i M^i)}=\exp{(-a_i L^i)}$. This happens to generate the $SO(3)$ fundamental representations, but this is not generally the case, it is ``accidental'' that the dimensions of the fundamental representation of $SO(3)$, $n=3$, and the dimension of its adjoint representation $n=3(3-1)/2=3$ is the same.

% Mention double cover here?

@@

The definition of the dual representation for a Lie algebra is as follows:\footnote{This definition also applies beyond finite-dimensional vector space with a basis since the transpose operation can be generalised beyond matrices.}
\df{If ${\mathfrak l}$ is a Lie algebra with a representation $\rho$ on a vector space $V$, then the {\bf dual representation} on the dual space $V^*$ is $\rho^*$ given by
\[ \rho^*(x) = -\rho(x)^T, \]
for all $x\in\mathfrak l$.}
To see why this makes sense we need to look at the exponential map of the dual representation. If $X_i$ is a generator in the representation $\rho$, then the dual representation $\rho^*$ tells us to use instead $-X_i^T$. In the exponentiation the group element $g=\exp(iX_i)$ becomes the group element $\exp(-iX_i^T)=\exp(-iX_i)^T=(g^{-1})^T$ with the dual representation.
% For this to be a group representation we must have that a product of two group elements $(g_1^{-1})^T(g_2^{-1})^T$

@@

>>>>

\subsubsection{$SO(2)$}
Given that $SO(2)$ has only one parameter, we can write a general group member $R$ as parameterised by $\theta$
\begin{equation}
R(\theta)=\left[\begin{matrix} \cos\theta &  -\sin\theta \\ \sin\theta&  \cos\theta \end{matrix}\right].
\label{eq:SO2_parameterisation}
\end{equation}
As expected, we recognise this as the matrix of rotations of an angle $\theta$ around a point in the plane, and it represents the (orientation preserving) symmetries of a circle. Some tinkering with this representation will show that $SO(2)$ is in fact an abelian group, despite the matrix definition. From a physical viewpoint this should be expected: the order of rotations in the plane should not matter.

@@

For matrix groups there is an alternative method to the generating functions for finding the generators that sees a lot of use. Since we are looking for infinitesimal deviations from the identity (matrix) $I$, we add an infinitesimal matrix $dM$ (the generator) and look at the properties of $dM$ given the definition of the group. Let us take $SO(2)$ as an example. Here a matrix in the group $O$ must fulfil $O^TO=I$ and $\det O=1$. Writing near the identity $O=I+dM$ we get the requirement $O^TO=I+dM+dM^T=I$ to first order in the infinitesimal. Thus the generator must fulfil $dM=-dM^T$ (it is a real, anti-symmetric matrix). This means that the diagonal of $dM$ must be zero, and in two dimensions, up to a constant, the only matrix that fits the role of the generator is
\[ X=\left[\begin{matrix} 0 & 1 \\ -1 & 0 \end{matrix}\right],\]
so that the group element can be written $O=I+d\theta X$, where $d\theta$ is some infinitesimal parameter of the group. To check the determinant we need to consider how we use the generator to recreate the group since the generator is defined through an infinitesimal change. For reasons that will become clearer later we use the exponential $O=\exp(\theta X)$, which to first order in an infinitesimal parameter $d\theta$ recreates the behaviour near the identity we used.
By the properties of matrix exponentiation we  can now check that $\det O=\det \exp(\theta X)=\exp(\theta \Tr X)=\exp(0)=1$. Since $X$ is an anti-Hermitian matrix, and we as physicists are more comfortable as Hermitian, we tend to use instead
\[ X=-i\left[\begin{matrix} 0 & 1 \\ -1 & 0 \end{matrix}\right]=\left[\begin{matrix} 0 & -i \\ i & 0 \end{matrix}\right],\]
as the generator, writing the group element as $O=\exp(i\theta X)$.

A similar calculation for $SO(3)$, which has three free parameters, gives the generators $X^i=\frac{1}{2}L^i$, where $L^i$ are the three matrices in Eq~(\ref{eq:SO3_generators}). These generators have the exact same commutator as the $SU(2)$ generators, $[X_i,X_j]=\epsilon_{ijk}X^k$, up to a constant factor, indicating a structural similarity between the groups $SU(2)$ and $SO(3)$. For $SO(3)$ we can further find a corresponding set of differential generators
\[
L_x=-i\left(y\frac{\partial}{\partial z}-z\frac{\partial}{\partial y} \right),\quad
L_y=-i\left(z\frac{\partial}{\partial x}-x\frac{\partial}{\partial z} \right),\quad
L_z=-i\left(x\frac{\partial}{\partial y}-y\frac{\partial}{\partial x} \right),
\]
see Ex.~\ref{ex:SO2_diff_generators}.
Up to a constant factor of $\hbar$ you may recognise these operators as the angular momentum operators in quantum mechanics. This is not a coincidence, the angular momentum operators in quantum mechanics are a representation of the generators of the group of rotations in three dimensions. In fact, the possibility of going back and forth between differential representations and matrix representations of the Lie group is the essence of the duality between the Schr\"odinger and Heisenberg pictures of quantum mechanics.

@@

Since Lie algebras are vector spaces, a {\bf subalgebra} $\mathfrak m$  of a Lie algebra $\mathfrak l$ is simply a subspace $\mathfrak m\subseteq \mathfrak l$. An {\bf ideal} is a subalgebra that satisfies $[\mathfrak l, \mathfrak m] \subseteq \mathfrak m$. A {\bf simple} Lie algebra is a non-abelian Lie algebra (non-zero structure constants)  without any proper ideals. A Lie algebra is {\bf semi-simple} if it is a direct sum of simple Lie algebras.

@@

From what we learnt in Section \ref{sec:Lie_groups} the generators $X_i$ of an $d$-dimensional Lie group then span a $d$-dimensional Lie algebra with the commutator as the Lie bracket.\footnote{If you are really paranoid here, you may question wether these generators are guaranteed to be linearly independent. However, since the Lie group is parameterised by $d$ parameters, if the generators were not linearly independent we should be able to remove one of the parameters, leading to a contradiction. Stop it!} Since this is also a vector space, any element $X$ of the algebra can be written in terms of the generators $X_i$, as $X=a_iX^i$. This is called the {\bf tangent space} of the group at the identity. The differences between choices in the definition of the generators and the normalisation of the structure constants disappear as they can be absorbed into a rescaling and linear combinations of the basis $\{X_i\}_{i=1}^d$, this makes the Lie algebra of the group {\it unique}.

@@

We saw in the previous subsection that -- just as for groups -- we have many different way of writing down a single Lie algebra. We then need a formal definition of representations of Lie algebras.
\df{A {\bf representation} $\pi$ {\bf of a Lie algebra} $\mathfrak l$ on a vector space $V$  is a map
\[\pi: \mathfrak{l} \to \mathfrak{gl}(V), \]
where $ \mathfrak{gl}$ is the Lie algebra of all linear maps $V\to V$ (endomorphisms) with its bracket defined as the composition $[r,s]\equiv r\circ s-s\circ r$, for all $r,s\in \mathfrak{gl}$, and where the map $\pi$ is homomorphic (structure preserving) for the Lie algebra bracket so that for all $x,y\in \mathfrak{l} $,
\[ \pi([x,y]) =\pi(x)\circ\pi(y)-\pi(y)\circ\pi(x). \]
}
Again this is an abstract definition that is actually simple than it looks. For finite dimensional vector spaces $\mathfrak{gl}$ is the Lie algebra of the corresponding general linear matrix group, just as the definition of representations of groups was a map to the general linear group $GL(n)$. This algebra of $GL(n)$, called $\mathfrak{gl}(n)$, can be written up in terms of $n\times n$ matrices and their commutators, and in this case the multiplication operation $\circ$ is just matrix multiplication in hiding. One word of warning here though, while the group $GL(n)$ contained invertible matrices, there is no guarantee that its generators are invertible matrices, {\it e.g.} we saw earlier that the generators of $SO(3)$ were the matrices $L_i$ in Eq~(\ref{eq:SO3_generators}) that are not invertible since $\det L_i=0$.

What makes this definition of representations even simpler for Lie algebras is that {\bf Ado's theorem}~\cite{Ado:1935} proves that every finite-dimensional Lie algebra has a faithful (isomorphic to the Lie algebra) representation on a finite-dimensional vector space.\footnote{For a modern version of a proof of Ado's theorem, see Terence Tao's blog \url{https://terrytao.wordpress.com/2011/05/10/ados-theorem/}.} In other words, the Lie algebras for the groups we most care about, Lie groups with a finite number of parameters and thus finite number of generators, is guaranteed to have representations using matrices.

As for groups, we can talk about irreducible representations, irreps, of the Lie algebras, which are representations such that $V$ has no proper invariant subspace under the representation, meaning that we cannot construct proper sub-representations.
% Maybe say something about similarity transforms between different Lie algebras 
An important result for representations of Lie groups is {\bf Weyl's complete reducibility theorem} which says that the any representation of a finite-dimensional semi-simple Lie algebra  on a finite-dimensional space is isomorphic to a direct sum of irreducible representations. This means that the classification of the representations of semi-simple Lie algebras can be done in a very systematic way. In particular, every semi-simple Lie algebra is a subalgebra of $\mathfrak {sl}$, the Lie algebra of the special linear group, and all our matrix groups have simple or semi-simple Lie algebras.
%TODO: extend to semi-simple representation theory here

In Sec.~\ref{sec:generators} we saw the two sets of generators of the Lie algebras $\mathfrak{su(2)}$ and $\mathfrak{so(3)}$, derived from the defining matrix representations of the groups that have different dimensions. Since this is the same algebra this shows two distinct representations of the same algebra.

%%%
\subsection{Exponential map}
\label{sec:expmap}
As we discussed in Section \ref{sec:Lie_groups}, the generators describe the local structure of the group near the identity element. We can now finally look at how the group (and matrix representation) is reconstructed from the algebra. For this we use what is called the {\bf exponential map}. 
\df{The {\bf exponential map} from a finite-dimensional Lie algebra $\mathfrak l$ of a group $G$ to $G$ is defined by $\exp:\mathfrak l\to G$, where for $X\in\mathfrak l$ we get the element $g\in G$ given by
\begin{equation}
g=\exp(iX)\equiv\sum_{n=0}^{\infty}\frac{(ia_iX^i)^n}{n!}.
\end{equation}
Here the $X_i$ are the generators of $G$ and thus members of the Lie algebra, and $X=a_iX^i$.
}
By Ado's theorem any (finite-dimensional) Lie algebra can be represented by matrices, in which case the infinite sum in this definition is nothing more than the formal series definition of the matrix exponential introduced in Sec.~\ref{sec:unitary_groups}. However, we can also use the exponential map for differential representations where the $X_i$ are differential operators, and for infinite-dimensional Lie groups, but the technical aspects why this extension works is a little beyond the scope of these notes.

@@

If we have two elements from the map $g=\exp{(iX)}$ and $g'=\exp{(iY)}$, then $gg'=\exp{(iX)}\exp{(iY)}$, and we would like to show that  $\exp{(iX)}\exp{(iY)}=\exp{(iZ)}$, where $Z\in\mathfrak l$. However, unless $X$ and $Y$ commute we cannot  easily join them in the same exponential. We here use the {\bf Baker--Campbell--Haussdorff  formula},
\begin{equation}
Z=X+Y+\frac{1}{2}[X,Y]+\frac{1}{12}[[X,Y],Y]-\frac{1}{12}[[X,Y],X]+\ldots,
\label{eq:BKH_matrices}
\end{equation}
where the omitted terms involve Lie brackets of four or more elements, and where every element in the series contains the commutator $[X,Y]$. If $X$ and $Y$ are close enough to the zero in $\mathfrak l$ (corresponding to the identity element in $G$) then this series can be shown to converge to an element in $\mathfrak l$.\footnote{This formula holds even if the elements in the Lie algebra ${\mathfrak l}$ are not represented as finite-dimensional matrices.} Thus the exponential constructs a group at least locally around the identity element.

@@

For $SU(2)$ the generators taken from the fundamental representation $\mathbf 2$ were proportional to the Pauli matrices $X_i= \frac{1}{2}\sigma_i$, and the exponential map generating a group member $U$ is thus, as we alluded to earlier, $U=e^{i\alpha_iX^i}$, where $\alpha_i$ are the parameters. For $SO(3)$ the (Hermitian) generators taken from the fundamental representation $\mathbf 3$ are the matrices $Y_i=-iL_i$, where the $L_i$ were given in (\ref{eq:SO3_generators}), and the exponential map is $O=e^{i\alpha_iY^i}$.
Since $SU(2)$ and  $SO(3)$ have the same algebra for $X_i$ and $Y_i$, but the groups are {\it not} isomorphic,  in what sense, if any, does these exponential maps generate different elements? The answer to that is that $SU(2)$ is a {\bf double cover} of $SO(3)$, it has double the number of elements. There exists a two-to-one group homomorphism $f:SU(2)\to SO(3)$. This can be represented by $U(\alpha_1,\alpha_2,\alpha_3)\to O(\alpha_1,\alpha_2,\alpha_3)$. To see that this is two-to-one, assume that $\alpha_1=\alpha_2=0$. Then $O(0,0,\alpha_3)=O(0,0,\alpha_3+2\pi)$ since a rotation by an extra $2\pi$ does nothing. For $SU(2)$ we instead have
\[
U(0,0,\alpha_3)=\left[\begin{matrix} e^{i\frac{\alpha_3}{2}} & 0 \\ 0 & -e^{i\frac{\alpha_3}{2}} \end{matrix}\right] \quad \text{and}\quad U(0,0,\alpha_3+2\pi)=\left[\begin{matrix} e^{i\frac{\alpha_3}{2}+i\pi} & 0 \\ 0 & -e^{i\frac{\alpha_3}{2}+i\pi} \end{matrix}\right]=-U(0,0,\alpha_3)
\]
Thus the $SU(2)$ matrices has a period $4\pi$ versus the period $2\pi$ for $SO(3)$.\footnote{The argument can be made more formal and independent of the parameterisation, but is to extensive to repeat here.} Notice that this means that the three-dimensional representation of the Lie algebra coming from the generators of $SO(3)$ does not generate all the elements of $SU(2)$, while the two-dimensional one does.

Beyond the fundamental representation $SU(2)$ has irreducible representations in all dimensions $\mathbf 3$, $\mathbf 4$, $\mathbf 5$, \ldots. This is not true in general, for example $SU(3)$ has irreducible representations only in some dimensions, starting with $\mathbf 3$, $\mathbf 6$, $\mathbf 8$, and $\mathbf{10}$, \ldots.

@@

Note that the dimension $n$ of the matrices in the adjoint representation for $\mathfrak so(m)$ and $\mathfrak su(m)$, which is equal to the number of independent parameters, is $n=m(m-1)/2$ and $n=m^2-1$, respectively. For the matrices in the representation derived from the defining representation of the group the dimensions is the same as the dimension of the defining matrices, $m$. This means that in most cases the matrices in the adjoint representation is larger, with the exception of  $\mathfrak{so}(2)$ and  $\mathfrak{so}(3)$.

We can now briefly return to the $SU(2)$ and $SO(3)$ examples. These groups have the same structure constants because they have the same commutators between generators. Thus the adjoint representations of their algebras are the same. With structure constants $C^{~~k}_{ij}=\epsilon_{ij}^{~~k}$ we get the elements of the adjoint matrices $(M_i)_{jk} =-i\epsilon_{ijk}$, which gives $M_i=-iL_i$, with the $L_i$ being the matrices in (\ref{eq:SO3_generators}).

From the adjoint representation of the Lie algebra we can construct  the adjoint representations of the Lie group. For example the adjoint representation of  $\mathfrak{su}(2)$ or $\mathfrak{so}(3)$ has the exponential map $\exp{(ia_i M^i)}=\exp{(-a_i L^i)}$. This happens to generate the $SO(3)$ fundamental representations, but this is not generally the case, it is ``accidental'' that the dimensions of the fundamental representation of $SO(3)$, $n=3$, and the dimension of its adjoint representation $n=3(3-1)/2=3$ is the same.

@@

The definition of the dual representation for a Lie algebra is as follows:\footnote{This definition also applies beyond finite-dimensional vector space with a basis since the transpose operation can be generalised beyond matrices.}
\df{If ${\mathfrak l}$ is a Lie algebra with a representation $\rho$ on a vector space $V$, then the {\bf dual representation} on the dual space $V^*$ is $\rho^*$ given by
\[ \rho^*(x) = -\rho(x)^T, \]
for all $x\in\mathfrak l$.}
To see why this makes sense we need to look at the exponential map of the dual representation. If $X_i$ is a generator in the representation $\rho$, then the dual representation $\rho^*$ tells us to use instead $-X_i^T$. In the exponentiation the group element $g=\exp(iX_i)$ becomes the group element $\exp(-iX_i^T)=\exp(-iX_i)^T=(g^{-1})^T$ with the dual representation. This is the only sensible representation that fulfils the group multiplication properties as matrices acting on row-vectors.
% For this to be a group representation we must have that a product of two group elements $(g_1^{-1})^T(g_2^{-1})^T$

@@

\begin{Exercise}[difficulty={1}]
Use the parameterisation of $SO(2)$ in (\ref{eq:SO2_parameterisation}) to find the generator of the group using generating functions and show that you get the same answer as when using the infinitesimal matrix demonstrated in the notes.
\end{Exercise}
